Mergeable Types

### 2017-02-27 paper/model.tex
Edits.

The merge operation for lists is composed of two separate functions,
\C{edit\_seq} and \C{op\_transform}. Both these functions have been
implemented in less than 50 lines of OCaml included in the appendix.
The functions are described below.
\begin{itemize}
  \item \C{edit\_seq} takes a pair of lists, \C{v} and $\C{v'}$, and
  computes the shortest sequence of list operations that need to be
  applied on \C{v} to obtain \C{v'}. Such a sequence is called an
  \emph{edit sequence}. The length of the sequence corresponds to the
  standard notion of the \emph{edit distance} between the two lists,
  which can be computed in polynomial time, for e.g., using the
  Wagner-Fisher algorithm~\cite{wagner-fischer}. \C{edit\_seq} uses a
  slight modification of Wagner-Fischer, that also lets it compute the
  edit sequence (an \C{edit list}). The implementation specified in
  Fig.~\ref{fig:mlist-sig} represents edits using the type \C{edit}.
  Constructors \C{I}, \C{D}, and \C{S} stand for \C{insert},
  \C{delete}, and \C{subst}, respectively\footnote{
    Note that \C{I}, \C{D} and \C{S} are \emph{not} effects; no list
    function generates them, and they are not exchanged between
    concurrent threads. They are simply tags used by the list merge
    operation for convenience.
  }. The subst constructor also carries the \C{A.t} element that was
  substituted. The element serves as the \C{lca} argument to
  \C{A.merge} in case of concurrent conflicting substitutions.
  Fig.~\ref{fig:list-eg} illustrates edit sequences for a sample list.
  The sequence \C{[I(c,0); S(3,c,s)]} maps the list \C{[a;b;c]} to
  \C{[c;a;b;s]} because \C{subst 3 s (insert c 0 [a;b;c]) =
  [c;a;b;s]}.
  \item  \C{op\_transform} takes a pair of edit sequences, $s_1$ and
  $s_2$, that map a list $v$ to two different lists, $v_1$ and $v_2$
  (e.g., Fig.~\ref{fig:list-eg}), and transforms $s_1$ to $s_1'$ such
  that $s_1'$ has the same effect on $v_2$ as $s_1$ had on $v$.  For
  instance, in Fig.~\ref{list-eg}, let $s_1$ be the edit sequence
  \C{[D(1); S(1,c,d)]}, which maps the list \C{[a;b;c]} ($v$) to
  \C{[a;d]} ($v_1$). The \C{D} edit deletes the first element of $v$.
  However, if $s_1$ is to be applied on the list \C{[c;a;b;s]}
  ($v_2$), which has a new element at position 0, the \C{D} edit must
  delete the the second element from $v_2$, if it were to have the
  same effect as it did on $v$. We say \C{D(1)} in $s_1$ has to be
  transformed w.r.t the sequence $s_2 = \C{[I(c,0);S(3,c,s)]}$.  The
  function \C{op\_transform} (Fig.~\ref{fig:mlist-sig}) computes such
  \emph{operational transformations} for an edit sequence $s_1$ w.r.t
  another edit sequence $s_2$. A notable transformation rule is for a
  substitution w.r.t another substitution at the same position. Since
  the substituting items could be different, the function relies on
  the item merge function (\C{A.merge}) to merge them into a single
  item. For instance, in Fig.~\ref{fig:list-eg}, there are
  simultaneous substitutions to the item \C{c} in $s_1$ and $s_2$;
  while $s_1$ substitutes it with \C{d}, $s_2$ does it with \C{s}. The
  merged list therefore substitutes \C{c} with the merge of \C{d} and
  \C{s}, defined as per the merge semantics of \C{A.t}.
\end{itemize}
The \C{MList.merge} can now be defined as a function that first
computes edit sequences, $s_1$ and $s_2$, for the two concurrent
lists, $v_1$ and $v_2$ w.r.t their common ancestor $v$, transforms one
of them (say, $s_1$) w.r.t $s_2$ to obtain $s_1'$, and finally applies
(the operations defined by) $s_1'$ to $v_2$ to obtain the merged list.

>>>>

The merge operation for lists is composed of two separate functions,
\C{edit\_seq} and \C{op\_transform}. Both these functions have been
implemented in less than 50 lines of OCaml included in the appendix.
The functions are described below.
\begin{itemize}
  \item \C{edit\_seq} takes a pair of lists, \C{v} and $\C{v'}$, and
  computes the shortest sequence of list operations that need to be
  applied on \C{v} to obtain \C{v'}. Such a sequence is called an
  \emph{edit sequence}. The length of the sequence corresponds to the
  standard notion of the \emph{edit distance} between the two lists,
  which can be computed in polynomial time~\cite{wagner-fischer}.
  \C{edit\_seq} modifies one such algorithm to also compute the edit
  sequence (an \C{edit list}). The implementation specified in
  Fig.~\ref{fig:mlist-sig} represents edits using the type \C{edit}.
  Constructors \C{I}, \C{D}, and \C{S} stand for \C{insert},
  \C{delete}, and \C{subst}, respectively\footnote{
    Note that \C{I}, \C{D} and \C{S} are \emph{not} effects; no list
    function generates them, and they are not exchanged between
    concurrent threads. They are simply tags used by the list merge
    operation for convenience.
  }. The subst constructor also carries the \C{A.t} element that was
  substituted. The element serves as the \C{lca} argument to
  \C{A.merge} in case of concurrent conflicting substitutions.
  Fig.~\ref{fig:list-eg} illustrates edit sequences for a sample list.
  The sequence \C{[I(c,0); S(3,c,s)]} maps the list \C{[a;b;c]} to
  \C{[c;a;b;s]} because \C{subst 3 s (insert c 0 [a;b;c]) =
  [c;a;b;s]}.
  \item  \C{op\_transform} takes a pair of edit sequences, $s_1$ and
  $s_2$, that map a list $v$ to two different lists, $v_1$ and $v_2$
  (e.g., Fig.~\ref{fig:list-eg}), and transforms $s_1$ to $s_1'$ such
  that $s_1'$ has the same effect on $v_2$ as $s_1$ had on $v$.  For
  instance, in Fig.~\ref{list-eg}, let $s_1$ be the edit sequence
  \C{[D(1); S(1,c,d)]}, which maps the list \C{[a;b;c]} ($v$) to
  \C{[a;d]} ($v_1$). The \C{D} edit deletes the first element of $v$.
  However, if $s_1$ is to be applied on the list \C{[c;a;b;s]}
  ($v_2$), which has a new element at position 0, the \C{D} edit must
  delete the the second element from $v_2$, if it were to have the
  same effect as it did on $v$. We say \C{D(1)} in $s_1$ has to be
  transformed w.r.t the sequence $s_2 = \C{[I(c,0);S(3,c,s)]}$.  The
  function \C{op\_transform} (Fig.~\ref{fig:mlist-sig}) computes such
  \emph{operational transformations} for an edit sequence $s_1$ w.r.t
  another edit sequence $s_2$. A notable transformation rule is for a
  substitution w.r.t another substitution at the same position. Since
  the substituting items could be different, the function relies on
  the item merge function (\C{A.merge}) to merge them into a single
  item. For instance, in Fig.~\ref{fig:list-eg}, there are
  simultaneous substitutions to the item \C{c} in $s_1$ and $s_2$;
  while $s_1$ substitutes it with \C{d}, $s_2$ does it with \C{s}. The
  merged list therefore substitutes \C{c} with the merge of \C{d} and
  \C{s}, defined as per the merge semantics of \C{A.t}.
\end{itemize}
The \C{MList.merge} can now be defined as a function that first
computes edit sequences, $s_1$ and $s_2$, for the two concurrent
lists, $v_1$ and $v_2$ w.r.t their common ancestor $v$, transforms one
of them (say, $s_1$) w.r.t $s_2$ to obtain $s_1'$, and finally applies
(the operations defined by) $s_1'$ to $v_2$ to obtain the merged list.
